---
title: "p8105_hw3_lg3158"
author: "Ling"
date: "10/17/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  out.width = "90%")
library(tidyverse)
library(p8105.datasets)
library(knitr)
```

# Problem 1

```{r}
data("instacart")
```

- A short description of the dataset:

  - The dataset `instacart` has `r ncol(instacart)` variables and `r nrow(instacart)` observations. The dataset includes `r nlevels(as_factor(pull(instacart, department)))` departments and `r nlevels(as_factor(pull(instacart, aisle)))` categories of products. The brief summary of this dataset is below.

```{r}
skimr::skim(instacart)
```


```{r}

instacart_summary = instacart %>%
  group_by(aisle) %>%
  summarise(items_sold = n()) %>%
  mutate( sell_rank = min_rank(desc(items_sold))) %>%
  arrange(sell_rank)

kable(instacart_summary[1:5, ], caption = "the top 5 most ordered aisles")
```

- From the summary and the table above, we can see that:
  - There are `r nlevels(as_factor(pull(instacart,aisle)))` aisles;
  - The fresh vegetable, fresh fruits, packaged vegetable fruits, yogurt, and packaged cheese are the aisles that most items are ordered from.
  
- Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly.

```{r echo = TRUE}
instacart_summary %>%
  filter(items_sold > 10000) %>%
  mutate(aisle = factor(aisle, levels = aisle[order(sell_rank, decreasing = T)])) %>%
  ggplot(aes(x = aisle, y = items_sold)) +
    geom_bar(stat = "identity") +
    labs( x = "the selling aisle",
          y = "items that are sold") +
    coord_flip()
```

- Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in the table.

```{r}
instacart_3pops = instacart %>%
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle ==  "packaged vegetables fruits") %>%
  select(product_name, aisle) %>%
  group_by(aisle, product_name) %>%
  summarise(items_sold = n()) %>%
  mutate(sell_rank = min_rank(desc(items_sold))) %>%
  arrange(aisle, sell_rank) %>%
  filter(sell_rank <= 3)

kable(instacart_3pops)
```

- Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.
```{r}
instacart_week = instacart %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  select(product_name, order_dow, order_hour_of_day) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean_order_time = mean(order_hour_of_day)) %>%
  mutate(order_dow = replace(order_dow, order_dow == 0, "Sunday"),
        order_dow = replace(order_dow, order_dow == 1, "Monday"), 
        order_dow = replace(order_dow, order_dow == 2, "Tuesday"), 
        order_dow = replace(order_dow, order_dow == 3, "Wednesday"),
        order_dow = replace(order_dow, order_dow == 4, "Thursday"),
        order_dow = replace(order_dow, order_dow == 5, "Friday"),
        order_dow = replace(order_dow, order_dow == 6, "Saturday")) %>%
  pivot_wider(names_from =  order_dow, values_from = mean_order_time)

kable(instacart_week)
```

# problem 2

- data importing and data cleaning
```{r}
data("brfss_smart2010") 
brfss = brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  rename(state = locationabbr,
         county = locationdesc) %>%
  mutate(county = substr(county, 6, 40),
         response = reorder(response, levels = c("poor", "Fair","Good", "Very Good", "Excellent"))) %>%
  arrange(response)
  
  summary(brfss)

```

